### 2018-01-03 toDelete/PositionPaper/lncs-pandoc-master/paper.tex
update files

\documentclass[]{llncs}

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments

% lncs 
\usepackage{makeidx}
\institute{FH Dortmund, \\ \texttt{}}

\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={Daniel Paredes},
            pdftitle={The power of dark silicon},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{5}

\title{The power of dark silicon}
\author{Daniel Paredes}
\date{}

\begin{document}
\maketitle
\begin{abstract}
Sit amet mauris. Curabitur a quam. Aliquam neque. Nam nunc nunc, lacinia
sed, varius quis, iaculis eget, ante. Nulla dictum justo eu lacus.
Phasellus sit amet quam. Nullam sodales. Cras non magna eu est
consectetuer faucibus. Donec tempor lobortis turpis. Sed tellus velit,
ullamcorper ac, fringilla vitae, sodales nec, purus. Morbi aliquet risus
in mi.

% lncs keyword extension 
\keywords{hope, luke, ewoks}

\end{abstract}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

I need an introduction\ldots{}

\hypertarget{using-argumentation-to-explain-ambiguity-in-requirements-elicitation-interviews}{%
\section{Using Argumentation to Explain Ambiguity in Requirements
Elicitation
Interviews}\label{using-argumentation-to-explain-ambiguity-in-requirements-elicitation-interviews}}

One of the major causes of ambiguities in elicitation interviews is the
presence of tacit knowledge. In some cases ambiguities can't be always
be explained as separated term, sometimes it is required to have a
context. Even under these circumstances, the analyst must be able to
identify and alleviate them in order to elicit relevant information of
the system. For this purpose it is necessary to provide the analyst with
proper tools. In this paper Yehia Elrakaiby et al. {[}1{]} proposed a
theoretical framework to overcome ambiguity during interviews in the
elicitation phase. The framework is based on the ``Argumentation
theory''.

In that sense, Elrakaiby et al.~focus on one type of ambiguities, the
``acceptance unclarity''. An acceptance unclarity occurs every time the
analyst is able to assign an interpretation or meaning to the speech
fragment of the stakeholder, the interpretation matches the intended
meaning of the stakeholder, but the interpretation is not acceptable or
justified. It could be either because it seems to be inaccurate to
comprehend the problem, or analysts identify inconsistencies with their
current understanding of the problem or domain knowledge. By using
argumentation theory framework, statements and ambiguities can be
characterized as ``arguments'' and ``attacks'' respectively.

Argumentation theory models a type of human dialog based on arguments
and conclusions. It makes explicit attacks between arguments and the
argumentation flow that leads to conclusions. A basic model in this
framework is a pair \((A,D)\), where \(A\) is a set of arguments and
\(D\) is a set of attacks among those arguments. For example, a set
\(A\) is defined as \(A = \{A1,A2,A3\}\), and a possible set of attacks
could be \(D = \{(A1,A3)\}\). Which means that if A1 is realizable then
A3 can't be realizable.

In the paper, Elrakaiby et al.~models statements given by the
stakeholders, analysts domain knowledge and analysts inferences as
arguments, and ambiguities between them as attacks. For example, let say
the analyst listens the following statement \emph{the professor will
upload the task description within three days} (A1), but the analyst
know (domain knowledge) that \emph{the professor may be on a business
meeting} (A2), so the analyst think (inference) that \emph{it may be
possible that it will be take longer to upload the task description}
(A3). In this scenario the set of attacks \(D\) is given by
\(D = \{(A1,A3)\}\). Thus, since there is an attack it is possible to
ask for clarifications or details.

The theoretical framework proposed by Elrakaiby et al.~allows analysts
to detect and minimize ambiguities during elicitation interviews, while
most of the methods that focus on ambiguities analyze written texts. On
the other hand, this framework focuses in more complex ambiguities that
cannot be view as single terms.

\hypertarget{effect-of-domain-knowledge-on-elicitation-effectiveness-an-internally-replicated-controlled-experiment}{%
\section{Effect of Domain Knowledge on Elicitation Effectiveness: An
Internally Replicated Controlled
Experiment}\label{effect-of-domain-knowledge-on-elicitation-effectiveness-an-internally-replicated-controlled-experiment}}

The effectiveness of elicitation interviews may be influenced by analyst
skills or characteristics. In these high intensive oral communication
scenario the analyst must be able to draw out relevant information and
needs from the stakeholders. It has been reported that the effectiveness
of the interviews has a direct relation with the domain knowledge of the
analyst. However, there are also studies suggesting that in some cases
the domain knowledge have negatives effects in the effectiveness of
interviews.

In this paper {[}2{]}, Aranda et al.~studied the influence of the
analyst domain knowledge on the effectiveness of elicitation interviews.
The main question they tried to answer was

\emph{Does analyst domain knowledge influence (either positively or
negatively) the effectiveness of the requirements elicitation activity?}

For this purpose, the authors performed the study in two stages. They
performed an initial baseline experiment with two domain problems, and
then they performed an internal replication with two other domain
problems. Furthermore, the authors divided the elicitation process in
two phases. The elicitation phase which is the actual interview with the
stakeholder, and the reporting phase in which the analyst understands
and documents the information gathered in the elicitation phase.

One remark of the study is that the authors chose students because of
their lack of experience in elicitation interviews, isolation of the
domain knowledge, and to analyze, in the internal replication, the
influence of the training in requirements engineering in elicitation
interviews. In the study participated post-graduated students as
interviewers and two professors as interviewees. The students should
make open interviews and elicit the information afterwards. Moreover,
for each domain problem the students were separated in two groups based
on their level of domain knowledge, \emph{domain-aware} and
\emph{domain-ignorant}. On the other hand, the effectiveness of the
elicitation's was based on the comparison between the number of
concepts, processes and requirements elicited by the students and the
benchmarks.

The results of the baseline experiment suggest that the domain knowledge
of the analysts has no significant influence in the effectiveness of the
elicitation interviews. However, the results also suggest that the
domain knowledge of the interviewees has statistically significant
influence. Supplementary, the results of the internal replication also
suggest that the domain knowledge of the stakeholders is more relevant
than the analysts', in term of effectiveness of the elicitation
interviews. Nevertheless, these results also show that the positive
effects of the training in requirements engineering of the interviewers,
with these effects being as relevant as interviewees' domain knowledge.

\hypertarget{requirements-elicitation-towards-the-unknown-unknowns}{%
\section{Requirements elicitation: Towards the Unknown
Unknowns}\label{requirements-elicitation-towards-the-unknown-unknowns}}

Elicitation in requirements engineering is still problematic because
missing or mistaken requirements are hard to elicit, and this produce
projects delays and have financial implications. In this paper, the
authors propose an elicitation review framework (ERF) in order to
explore the different challenges related to the ``unknowness'' of the
domain knowledge. Therefore, proposing a road map of research to tackle
the different levels of the ``unknowness''.

The authors made use of the properties of \emph{expressible} (know
knowledge), \emph{articulated} (documented domain knowledge),
\emph{accessible} (need memory recall), \emph{relevant} (to the
project). With these four properties, authors definied the following
levels of unknowness in domain knowledge.

\begin{itemize}
\tightlist
\item
  Known knowns: knowledge that is expressible, articulated, and
  relevant.
\item
  Known unknowns: knowledge that is not expressible or articulated, but
  accessible and potentially relevant.
\item
  Unknown knowns: knownledge that is potentially accessible but not
  articulated.
\item
  Unknown unknowns: knownledge that is not expressible, articulated nor
  accessible but may be relevant.
\end{itemize}

This levels of knownledge implies problems of different perspective of
the elicitation process that involves analysts and stakeholders. The
\emph{known knowns} is a simple scenario since the counterparts are
aware of the domain problem which is expressable and articulated. In the
case of the \emph{known unknowns}, the analysts has a domain knowledge
and they should be able draw out the information from the stakeholders
that they might be unaware of. In contrast, in the \emph{unknow knowns}
scenario the stakeholders are the ones that are aware of the domain
knowledge but for some reason it is no articulated, and therfore, the
analysts once they notice any glimpse of the stakeholders information
they should able to discover and elicit that. On the other hand, in the
case of \emph{unknows unknowns} both counterparts are unaware of the
missing information that might be relevant to the system.

Based on the perspectives described before, the authors identified three
challenges to requirement elicitation. The first one is to identifying
tacit knowledge (unknown knowns), even in the case the analyst presume
it exist (known unknowns). The second challenge is that the analyst
should know what is relevant and should be articulated. And the last one
is the articulation of the knowledge. Thus, in the worst case scenario
the goal is go from the unknow unknows perspective to know knows.

Finally, the authors suggest four research directions to deal with the
unknowness problem. Starting with the known unknown in which the analyst
should challenging assumptions, implications, and relaxing domain
constrains in order to increase the probability to elicit the unaware
knownledge of the stakeholders. In the same direction, in the design
discovery, which is a variant of the latter case, the analyst should
deal with the the following statement ``I'll know what I want when I see
it''. Thus, there is a need of prototyping or simulations as part of the
elicitation process. In the case of the unknown knowns in which the
analyst can elicit tacit knownledge by taking into account the
background (political, cultural, emotional) of stakeholders, and use
that background as an ``emotional guidance'' in order to draw out the
relevant information. Finally, in the case of unknown unknows the
authors suggest to main approaches. The first one is based on the
``over-the-horizon'' knowledge, in which authors proposed a
socio-technical approach based on social media and e-communities. And
the other one is using counter examples that can challenges the
boundaries of the idea under development.

\hypertarget{what-ill-do}{%
\section{What i'll do}\label{what-ill-do}}

Explain acceptance unclarity Explain Dung's framework Explain ASPIC+
Inconsistencies It's highly influenced by the domain knowledge of the
analyst

\hypertarget{my-idea}{%
\subsection{My idea}\label{my-idea}}

knowledge domain has influence analyst domain knowledge is statistically
relevant analyst training is even more relevant influence of tacit
knowledge in interviews

ambiguity in interviews - focus on interviewer must have domain
knowledge what i want to do is : possible benefits in requirement
elicitation is using two techniques based on the domain knowledge of the
analyst, sometimes it is good to have an expert and an ignorant of a
topic .

limitations of the technique based on domain knowledge and tacit
knowledge. how types of tacit knowledge may influence in the proposal
theory how domain knowledge may influence in the proposal theory

As an intro: analyst domain knowledge is statistically relevant analyst
training is even more relevant focus on the technique from the
perspective of knowledge and technique future of that technique to solve
more complex problems, from the prospective of knowledge and technique

\newpage

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-elrakaiby2017using}{}%
{[}1{]} Y. Elrakaiby, A. Ferrari, P. Spoletini, S. Gnesi, and B.
Nuseibeh, ``Using argumentation to explain ambiguity in requirements
elicitation interviews,'' in \emph{Requirements engineering conference
(re), 2017 ieee 25th international}, 2017, pp. 51--60.

\leavevmode\hypertarget{ref-aranda2016effect}{}%
{[}2{]} A. M. Aranda, O. Dieste, and N. Juristo, ``Effect of domain
knowledge on elicitation effectiveness: An internally replicated
controlled experiment,'' \emph{IEEE Transactions on Software
Engineering}, vol. 42, no. 5, pp. 427--451, 2016.

\end{document}