### 2018-01-07 toDelete/PositionPaper/lncs-pandoc-master/paper.tex
update files

\documentclass[]{llncs}

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments

% lncs 
\usepackage{makeidx}
\institute{FH Dortmund, \\ \texttt{}}

\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage{longtable,booktabs}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={Daniel Paredes},
            pdftitle={Influence of tacit knowledge in elicitation interviews},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{5}

\title{Influence of tacit knowledge in elicitation interviews}
\author{Daniel Paredes}
\date{}

\begin{document}
\maketitle
\begin{abstract}
Interviews are the most common technique in requirements elicitation.
However, the presence of tacit knowledge have an important influence on
interviews success. In this paper, we describe a tacit knowledge
classification. Then, we make use, with examples, of a theoretical
framework to detect and minimize the effects of tacit knowledge.
Finally, we analyze how likely is to have a successful elicitation
interviews based on the tacit knowledge classification.

% lncs keyword extension 
\keywords{interview, tacit knowledge, elicitation}

\end{abstract}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Requirements engineering is the subset of systems engineering concerned
with discovering, developing, tracing, analyzing, qualifying,
communicating and managing requirements that define the system at
successive levels of abstraction {[}1{]}. The first step of requirements
engineering is requirements elicitation which is defined as the process
of finding and formulation requirements {[}2{]}, and this process has
many barriers due to stakeholders are not able to express their needs,
explain what they do and why, conflicting demands, or new demands once
other are met, just to mention some examples.

One of the most used techniques to elicit requirements is open
interviews, and also is considered the most effective way to transfer
domain knowledge {[}3{]}. However, the success of interviews depends on
mainly two factors, analyst set of interviewing skills and stakeholders
domain knowledge; nonetheless, statistically analyst domain knowledge is
important but not as relevant as the other two{[}4{]}. Therefore,
analyst depends on the use of tools and techniques to detect and elicit
tacit knowledge form stakeholders. Tacit knowledge in requirements
engineering is defined as the knowledge that a stakeholders has, but
does not pass to the requirements analyst {[}5{]}.

In this paper we describe a classification of tacit knowledge in
elicitation interviews based on {[}6{]}. Then we apply the theoretical
framework described in {[}3{]} to examples to show how to deal with
tacit knowledge. The examples were selected in order to analyze each
scenario of the classification. Finally, we use the result from {[}4{]}
to identify in which scenario is most likely to perform a successful
interview.

The remainder paper is structured as follows. In section 2 we provide a
summary of the papers used to write this paper. In section 3 we present
our approach. Finally, in section 4, we present our conclusions.

\hypertarget{background}{%
\section{Background}\label{background}}

\hypertarget{requirements-elicitation-towards-the-unknown-unknowns-sutcliffe2013requirements}{%
\subsection{Requirements elicitation: Towards the Unknown Unknowns
{[}6{]}}\label{requirements-elicitation-towards-the-unknown-unknowns-sutcliffe2013requirements}}

Elicitation in requirements engineering is still problematic because
missing or mistaken requirements are hard to elicit, and this produce
projects delays and have financial implications. In this paper, the
authors propose an elicitation review framework (ERF) in order to
explore the different challenges related to the ``unknowness'' of the
domain knowledge. Therefore, proposing a road map of research to tackle
the different levels of the ``unknowness''.

The authors made use of the properties of \emph{expressible} (known
knowledge), \emph{articulated} (documented domain knowledge),
\emph{accessible} (need memory recall), \emph{relevant} (to the
project). With these four properties, authors defined the following
levels of unknowness in domain knowledge.

\begin{itemize}
\tightlist
\item
  Known knowns: knowledge that is expressible, articulated, and
  relevant.
\item
  Known unknowns: knowledge that is not expressible or articulated, but
  accessible and potentially relevant.
\item
  Unknown knowns: knowledge that is potentially accessible but not
  articulated.
\item
  Unknown unknowns: knowledge that is not expressible, articulated nor
  accessible but may be relevant.
\end{itemize}

Based on the perspectives described before, the authors identified three
challenges to requirement elicitation. The first one is to identifying
tacit knowledge (unknown knowns), even in the case the analyst presume
it exist (known unknowns). The second challenge is that the analyst
should know what is relevant and should be articulated. And the last one
is the articulation of the knowledge. Thus, in the worst case scenario
the goal is go from the unknown unknowns perspective to known knowns.

Finally, the authors suggest four research directions to deal with the
unknowness problem. First, in \emph{known unknown} the analyst should
challenging assumptions, implications, and relaxing domain constrains in
order to increase the probability to elicit the unaware knowledge of the
stakeholders. In the same direction, the design discovery, which is a
variant of the latter case, the analyst should deal with the the
following statement ``I'll know what I want when I see it''. Thus, there
is a need of prototyping or simulations as part of the elicitation
process. In the case of the \emph{unknown knowns} the analyst can elicit
tacit knowledge by taking into account the background (political,
cultural, emotional) of stakeholders, and use that background as an
``emotional guidance'' in order to draw out the relevant information.
Finally, in order to deal with \emph{unknown unknowns} the authors
suggest two main approaches. The first one is based on the
``over-the-horizon'' knowledge which is a socio-technical approach based
on social media and e-communities. And the other one is using counter
examples that can challenges the boundaries of the idea under
development.

\hypertarget{using-argumentation-to-explain-ambiguity-in-requirements-elicitation-interviews-elrakaiby2017using}{%
\subsection{Using Argumentation to Explain Ambiguity in Requirements
Elicitation Interviews
{[}3{]}}\label{using-argumentation-to-explain-ambiguity-in-requirements-elicitation-interviews-elrakaiby2017using}}

One of the major causes of ambiguities in elicitation interviews is the
presence of tacit knowledge. In some cases ambiguities can't be always
be explained as separated term, sometimes it is required to have a
context. Even under these circumstances, the analyst must be able to
identify and alleviate them in order to elicit relevant information of
the system. For this purpose it is necessary to provide the analyst with
proper tools. In this paper Yehia Elrakaiby et al.~proposed a
theoretical framework to overcome ambiguity during interviews in the
elicitation phase. The framework is based on the ``Argumentation
theory''.

In that sense, the authors focus on one type of ambiguities, the
``acceptance unclarity''. An acceptance unclarity occurs every time the
analyst is able to assign an interpretation or meaning to the speech
fragment of the stakeholder, the interpretation matches the intended
meaning of the stakeholder, but the interpretation is not acceptable or
justified. It could be either because it seems to be inaccurate to
comprehend the problem, or analysts identify inconsistencies with their
current understanding of the problem or domain knowledge. By using
argumentation theory framework, statements and ambiguities can be
characterized as ``arguments'' and ``attacks'' respectively.

Argumentation theory models a type of human dialog based on arguments
and conclusions. It makes explicit attacks between arguments and the
argumentation flow that leads to conclusions. A basic model in this
framework is a pair \((A,D)\), where \(A\) is a set of arguments and
\(D\) is a set of attacks among those arguments. For example, a set
\(A\) could be defined as \(A = \{A1,A2,A3\}\), and a possible set of
attacks could be \(D = \{(A1,A3)\}\). Which means that if A1 is
realizable then A3 can't be realizable.

In the paper, statements given by the stakeholders, analysts domain
knowledge and analysts inferences are modeled as arguments, and
ambiguities between them as attacks. For example, let say the analyst
listens the following statement \emph{the professor will upload the task
description within three days} (A1), but the analyst know (domain
knowledge) that \emph{the professor may be on a business meeting} (A2),
so the analyst think (inference) that \emph{it may be possible that it
will be take longer to upload the task description} (A3). In this
scenario the set of attacks \(D\) is given by \(D = \{(A1,A3)\}\). Thus,
since there is an attack it is possible to ask for clarifications.

The theoretical framework proposed allows analysts to detect and
minimize ambiguities during elicitation interviews, while most of the
methods focus on ambiguities analysis in written texts. On the other
hand, this framework focuses in more complex ambiguities that cannot be
view as single terms.

\hypertarget{effect-of-domain-knowledge-on-elicitation-effectiveness-an-internally-replicated-controlled-experiment-aranda2016effect}{%
\subsection{Effect of Domain Knowledge on Elicitation Effectiveness: An
Internally Replicated Controlled Experiment
{[}4{]}}\label{effect-of-domain-knowledge-on-elicitation-effectiveness-an-internally-replicated-controlled-experiment-aranda2016effect}}

The effectiveness of elicitation interviews may be influenced by analyst
skills or characteristics. In this high intensive oral communication
scenario the analyst must be able to draw out relevant information and
needs from the stakeholders. It has been reported that the effectiveness
of the interviews has a direct relation with the domain knowledge of the
analyst. However, there are also studies suggesting that in some cases
the domain knowledge have negatives effects in the effectiveness of
interviews.

In this paper Aranda et al.~studied the influence of the analyst domain
knowledge on the effectiveness of elicitation interviews. The main
question they tried to answer was

\begin{quote}
\emph{Does analyst domain knowledge influence (either positively or
negatively) the effectiveness of the requirements elicitation activity?}
\end{quote}

For this purpose, the authors performed the study in two stages. They
performed an initial baseline experiment with two domain problems, and
then they performed an internal replication with two other domain
problems. One remark of the study is that the authors chose students
because of their lack of experience in elicitation interviews, isolation
of the domain knowledge, and to analyze, in the internal replication,
the influence of the training in requirements engineering in elicitation
interviews. In the study participated post-graduated students as
interviewers and two professors as interviewees. The students should
make open interviews and elicit the information afterwards. Moreover,
for each domain problem the students were separated in two groups based
on their level of domain knowledge, \emph{domain-aware} and
\emph{domain-ignorant}. On the other hand, the effectiveness of the
elicitation's was based on the comparison between the number of
concepts, processes and requirements elicited by the students and the
benchmarks.

The results of the baseline experiment suggest that the domain knowledge
of the analysts has no significant influence in the effectiveness of the
elicitation interviews. However, the results also suggest that the
domain knowledge of the interviewees has statistically significant
influence. Supplementary, the results of the internal replication also
suggest that the domain knowledge of the stakeholders is more relevant
than the analysts', in terms of effectiveness of elicitation interviews.
Nevertheless, these results also show that the positive effects of the
training in requirements engineering of the interviewers, with these
effects being as relevant as interviewees' domain knowledge.

\hypertarget{tacit-knowledge-in-elicitation-interviews}{%
\section{Tacit knowledge in elicitation
interviews}\label{tacit-knowledge-in-elicitation-interviews}}

\hypertarget{types-of-tacit-knowledge}{%
\subsection{Types of tacit knowledge}\label{types-of-tacit-knowledge}}

Many analyst consider interviews the most important elicitation
technique, and usually elicitation process starts with interviewing the
stakeholders. Interviews allow analysts to check their understanding
about the problem domain immediately and ask for clarifications.
Moreover, this technique is versatile in the sense of it is possible
that new and unexpected issues come up and they can be attacked at that
moment.

However, elicit tacit knowledge in interviews is still a hard task to
do. This tacit knowledge can be classified in four groups as describe in
{[}6{]}. The \textbf{known knowns}, \textbf{known unknowns},
\textbf{unknown knowns} and \textbf{unknown unknowns}. The known knowns
is the knowledge that is documented, expressible, and relevant to the
system. In order words, is the scenario in which it is possible to
elicit since the beginning all the requirements, there is no tacit
knowledge. The known unknowns is the knowledge that cannot be
expressible because the stakeholders are not aware of their lack of
domain knowledge, their might forget it. In this case analyst are aware
of the domain knowledge and therefore, their job is to challenge
assumptions or implications in order to elicit information that
stakeholders are unaware of. The unknown knowns is the knowledge that
stakeholders hold but it is not documented by some reason. Thus,
analysts should use their interviewing skill to elicit when they spot
any glimpse of that knowledge. The unknown unknowns is the knowledge
that neither the analysts nor stakeholders are aware but might be
relevant to the project.

\hypertarget{how-to-deal-with-tacit-knowledge}{%
\subsection{How to deal with tacit
knowledge}\label{how-to-deal-with-tacit-knowledge}}

The unknowness clarification presented in above suggest the presence of
three problems. The goal is to go from any scenario in which tacit
knowledge exists to known knowns. In order to achieve that goal and take
the advantages presented in elicitation interviews (immediately
clarification of assumptions and debate of implications) we are going to
use the theoretical framework described by Elrakaiby {[}3{]}. In this
framework we model the interview as a set of arguments and attacks
between them. In that sense, we will able to spot ambiguities and tacit
knowledge in terms of inconsistencies and insufficiency of information.
We will adapt some examples from {[}3{]} to show how each tacit
knowledge scenario can be addresed.

\textbf{known unknowns}: In the following example the domain knowledge
of the analyst is used to detect inconsistencies (attacks), because the
stakeholder didn't take into account relevant information about domain
application (location of trash bins).

A customers wanted a recycling-support system and one of the goals was
\emph{to avoid fines from the municipality for incorrect recycling}.
According to the domain knowledge of the analyst, \emph{trash bins are
placed along the streets}. Thus, the goal was inconsistent with the
domain knowledge of the analyst because \emph{there is no way to trace
back person who brakes the recycling rule}. This scenario can be modeled
using arguments and attacks as follows:

The arguments: (A1)trash bins are placed along the streets, (A2)since
trash bins are placed along the streets, the trash owner cannot be
traced back. (B1)people who do not recycle should be fined (B2)to fine
people, the municipality must be able to trace products in trash bins
back to their owners.

Attacks: A2 attacks B1, and vice versa.

\textbf{unknown knowns}: In the following example analyst challenges the
stakeholder statement, as a result the stakeholder reveals new
information (reply).

A customers wants to develop a system to allow patients to measure the
amount of glucose in their blood, and then send the result to their
general practitioner. If the glucose level is above a certain threshold,
the practitioner pays a visit to the patient. The customer said:
\emph{in one-two days the doctor sees the notification in the system}.
Then analyst asked: \emph{What happens if the doctor is not available
for any reason?}. The customer replied: \emph{The general practitioner
is substituted by another doctor who accesses the same system}.

The arguments: (A1)in one-two days the doctor sees the
notification,(B1)the doctor might be unavailable,(B2)since the the
doctor might be unavailable, the notification might be delayed.

The attacks: A1 attacks B2, and vice versa

\textbf{unknown unknowns}: In the following example the stakeholder
didn't notice the inconsistency between his statements, but the analyst
did and asked for clarification. However, the stakeholder couldn't give
an answers due to his ignorance.

One of the customers wants to develop an app to manage medical-related
reservation in Tuscani. The costumer mentioned that the \emph{current
reservation system, based on phone calls, was centralized}. During the
interview, the customer said that \emph{depending on where the
examination/visit will be, the patient has to call Nottola or Siena}.
The analyst could not understand how this realization was possible since
it means the reservation system is not centralized. Therefore the
analyst asked for clarification and the stakeholder couldn't give any
acceptable answer.

The arguments: (A1)the current reservation is centralized, (B1)depending
on where the examination/visit will be (B2)the patience has to call
Nottola or Siena, (B3) since it is possible to call either Nottola or
Siena the system is not centralized.

The attacks: A1 attacks B3, and vice versa

\hypertarget{limits-of-elicited-knowledge}{%
\subsection{Limits of elicited
knowledge}\label{limits-of-elicited-knowledge}}

In the last section we present how to use a theoretical framework based
on arguments and attacks to addressed the different levels of tacit
knowledge. In this section we will compare in which scenario of tacit
knowledge the elicitation interview would be more effective based on
Aranda et al.~work {[}4{]} about the effect of domain knowledge on
elicitation process.\\
Effectiveness is described in {[}4{]} as the total percentage of problem
domain elements identified compared to a benchmark. However, since we
have no access to any requirement data, we will use the results from
{[}4{]} to compare the different scenarios of tacit knowledge (known
unknowns, unknown knowns and unknowns unknowns).

The works presented in {[}4{]} concludes that: (1)analyst's domain
knowledge(ADK) has small but statistically significant effect on the
effectiveness, (2)stakeholders' domain knowledge(SDK) has a big positive
and significant influence and (3)analyst's interviewing skill and
general training in requirement activities have also a big positive and
significant influence. In this paper, we will focus only in comparing
based on the domain knowledge holder since that knowledge can be
measured.

\begin{longtable}[]{@{}lcc@{}}
\caption{Influence of domain knowledge in the different tacit knowledge
scenarios}\tabularnewline
\toprule
scenario & ADK & SDK\tabularnewline
\midrule
\endfirsthead
\toprule
scenario & ADK & SDK\tabularnewline
\midrule
\endhead
known unknowns & \(+\) & 0\tabularnewline
unknown knowns & 0 & \(++\)\tabularnewline
unknown unknowns & 0 & 0\tabularnewline
\bottomrule
\end{longtable}

We can observed in Table 1 that it is more likely to be effective in the
scenario of \emph{unknown knowns} because the stakeholders' domain
knowledge have more influence than analyst's domain knowledge. On the
other hand, the analyst's interviewing skills will have different
approaches in the case of \emph{known unknowns} and \emph{unknown
knowns}. In the first case, the interview will be more focused on
challenging assumptions and implications, and in the latter case on
spotting glimpses of stakeholders' hidden knowledge.

\hypertarget{conclusions}{%
\section{Conclusions}\label{conclusions}}

Elicitation interviews is not an easy task. Like in any intensive oral
communication there are misunderstandings related to tacit knowledge. In
our work we focus on analyzing the different levels of how tacit
knowledge{[}6{]} would affect the effectiveness of elicitation
interviews. Moreover, we use the theoretical framework described in
{[}3{]} to identify and manage tacit knowledge during interviews. And
finally, we analyze based on the results of {[}4{]} in which scenario of
tacit knowledge it is more likely to have a successful elicitation
interview.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-dick2017requirements}{}%
{[}1{]} J. Dick, E. Hull, and K. Jackson, \emph{Requirements
engineering}. Springer International Publishing, 2017.

\leavevmode\hypertarget{ref-lauesen2001softwarereq}{}%
{[}2{]} S. Lauesen, \emph{Software requirements: Styles and techniques},
1st ed. Pearson Education, 2001.

\leavevmode\hypertarget{ref-elrakaiby2017using}{}%
{[}3{]} Y. Elrakaiby, A. Ferrari, P. Spoletini, S. Gnesi, and B.
Nuseibeh, ``Using argumentation to explain ambiguity in requirements
elicitation interviews,'' in \emph{Requirements engineering conference
(re), 2017 ieee 25th international}, 2017, pp. 51--60.

\leavevmode\hypertarget{ref-aranda2016effect}{}%
{[}4{]} A. M. Aranda, O. Dieste, and N. Juristo, ``Effect of domain
knowledge on elicitation effectiveness: An internally replicated
controlled experiment,'' \emph{IEEE Transactions on Software
Engineering}, vol. 42, no. 5, pp. 427--451, 2016.

\leavevmode\hypertarget{ref-ferrari2015ambiguity}{}%
{[}5{]} A. Ferrari, P. Spoletini, and S. Gnesi, ``Ambiguity as a
resource to disclose tacit knowledge,'' in \emph{Requirements
engineering conference (re), 2015 ieee 23rd international}, 2015, pp.
26--35.

\leavevmode\hypertarget{ref-sutcliffe2013requirements}{}%
{[}6{]} A. Sutcliffe and P. Sawyer, ``Requirements elicitation: Towards
the unknown unknowns,'' in \emph{Requirements engineering conference
(re), 2013 21st ieee international}, 2013, pp. 92--104.

\end{document}